A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-16 papers/static-checker/ecoop2024/ecoop2024.tex
general revision

While designers can create tokens that behave in any way they want, it
is advisable to implement tokens that adhere to established standards.
This adherence guarantees interoperability and compatibility across
different platforms. A token standard defines a set of rules and an
interface that smart contracts must follow to ensure compatibility
with decentralized exchanges, wallets, and other tools within the
ecosystem. For example, ERC-20 \cite{erc} is a token standard used in Ethereum.

The Tezos blockchain has its own set of standards known as Financial
Applications (FA). Two notable standards are: (1) FA 1.2 \cite{fatzip}, designed for
fungible tokens, and (2) FA 2 \cite{fatzips}, a more versatile standard accommodating
both fungible and non-fungible tokens (NFTs). The standard outlines a
smart contract that functions as a ledger, associating identities with
balances. This ledger is designed to facilitate token transfer
operations and enable approvals for spending tokens from other
accounts. Any contract aiming to implement the standard must
incorporate the functions specified in
Listing~\ref{lst:fa12-interface} such as token transfer (transactions
between users), approval (allowing one user to send tokens from
another user's address),  allowance (browsing the amount of tokens
available for an account to sending from another address), user
balance (checking the balance of a user), and current supply (checking
the total supply of tokens across all addresses), where
\lstinline/view a r = pair a (contract r)/ implies an view entrypoint.
A view is an entrypoint that takes an
argument of type \lstinline/a/ and a callback that results in a value
of type \lstinline/r/. The callback in this case will be another
contract address. Further elaboration on the details of FA 1.2 can be
found in \cite{usdtz}, and we will delve into
these functions in subsequent 
sections.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
   storage :=  Pair (ledger: big_map address 
      (pair nat (map address nat))) 
      (Pair (admin: address) 
            (Pair (paused: bool) (total_supply: nat)))
   entrypoint %transfer
      parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
      pre-condition := not paused 
        && get_big_map (from, ledger) = 
           Some (v: pair nat (map address nat))
        && (v = Pair (n: nat) (m: map address nat))
        && get_map (Sender, m) = Some (k: nat) 
        && (n >= value) && (k >= value)
      post-condition := get_big_map (from, post_ledger) = 
           Some (post_v: option (pair nat (map address nat))) 
        && post_v = Pair (post_n: nat) (post_m: map address nat) 
        && get_map (Sender, post_m) = Some (post_k: nat) 
        && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}

@@

The crucial property to ensure the correctness of the \texttt{approve}
entrypoint is that changing the allowance value from a non-zero value
to another non-zero value is forbidden to prevent a corresponding
attack vector \cite{attack-vector}. The basic idea of this attack is
that when a non-zero value update is allowed, the \lstinline/spender/
can attempt to spend the entire allowance before the new amount is
updated and then spend again with the new value, resulting in a larger
amount than actually allowed. Hence, the contract call must fail when
both the parameter \lstinline/value/ is non-zero and the allowance of
\lstinline/spender/ is also non-zero. We verify this property by
generating a fail condition, so that a non-zero value update is
recorded as a fail condition, as shown in
Listing~\ref{lst:approve-udstz-contract}. This condition indicates
that the implementation ensures the non-zero value update prevention. 
\begin{lstlisting}[float=tp,captionpos=b,caption={Fail conditions for the \lstinline/approve/ entrypoint},label={lst:approve-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (Sender, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (spender, sbvar2) = Some sbvar3 &&
   not (value = 0) &&
   not (sbvar3 = 0): 'UnsafeAllowanceChange'
\end{lstlisting}
If the update is allowed, then the updated value needs to be recorded
correctly in the storage. This property can be divided into two
subproperties, which can be checked separately. Property 1: if the
parameter \lstinline/value/ is zero, then the 
corresponding allowance value of the \lstinline/spender/ from the
owner (which is \SENDER\ in this call) should be zero. This
subproperty is specified as shown in
Listing~\ref{lst:specification-approve}. Property 2: if
\lstinline/value/ is non-zero and the allowance of the \lstinline/spender/ is zero, then the \lstinline/value/ should be recorded in the storage after the call. We specify this property using pre- and post-conditions, as illustrated in Listing~\ref{lst:---approve---to-be-provided}.

\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/approve/ entrypoint},label={lst:specification-approve},numbers=left]
entrpoint %approve
   parameter := Pair (spender: address) (value: nat)
   pre-condition := not (paused = true)
      && get_big_map (Sender, ledger) = 
        Some (v: pair nat (map address nat))
      && v = Pair (n: nat) (m: map address nat) 
      && get_map (spender, m) = Some (k: nat) 
      && not (k = 0) | (value = 0)
   post-condition := get_big_map (Sender, post_ledger) = 
        Some (post_v:  (pair nat (map address nat))
      && post_v = Pair (post_n: nat) (post_m: map address nat) 
      && get_map (spender, post_m) = Some (post_k: nat) 
      && (post_k = 0) ;

   pre-condition := not (paused = true) 
      && get_big_map (Sender, ledger) = 
        Some (v: option (pair nat (map address nat))) 
      && v = Pair (n: nat) (m: map address nat) 
      && get_map (spender, m) = Some (k: nat) 
      && (k = 0) | not (value = 0)
   post-condition := get_big_map (Sender, post_ledger) = 
        Some (post_v: option (pair nat (map address nat))) 
      && post_v = Pair (post_n: nat) (post_m: map address nat) 
      && get_map (spender, post_m) = Some (post_k: nat) 
      && (post_k = value) ;
\end{lstlisting}

@@

As an example, we demonstrate how to verify the \lstinline/getBalance/
entrypoint,  shown in Listing~\ref{lst:specification-get-balance}. The
other entrypoints \lstinline/getAllowance/ and
\lstinline/getTotalSupply/ can be specified similarly.  There are two
cases to consider: (1) if the \lstinline/owner/ is in the
\lstinline/ledger/ big map and (2) if it is not. For the former case,
lines 5-8 specify that \lstinline/owner/ is in the \lstinline/ledger/
big map, as the \lstinline/Get_big_map/ function returns a value
\lstinline/Some value/ of the \TOPTION\ type, and this
\lstinline/value/ is a pair of a natural number \lstinline/k/ and a
map \lstinline/m/. The postcondition guarantees that the storage
remains unchanged, as specified in lines 10-13, and the call returns a
\lstinline/TRANSFER_TOKENS/ operation with the \lstinline/k/ number
and \lstinline/Amount/ to be passed to address \lstinline/c/. The
precondition in lines 16-17 specifies the latter case when the
\lstinline/owner/ does not exist in the big map \lstinline/ledger/; in
this case, the number \lstinline/0/ should be sent, and again, the
storage should remain unchanged.

\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/getBalance/ entrypoint},label={lst:specification-get-balance},numbers=left]
entrypoint %getBalance
   parameter := Pair (owner: address) (c: contract nat)
   pre-condition :=  get_big_map (owner, ledger) = 
        Some v: option (pair nat (map address nat)) 
      && v  = Pair (k: nat) (m: (map address nat))
   post-condition := post_ledger = ledger 
      && post_admin = admin 
      && post_paused = paused 
      && post_total_supply = total_supply 
      && post_Balance = Balance 
      && transfer_token (k, Amount, c) ;

   pre-condition :=  get_big_map (owner, ledger) = 
        None option (pair nat (map address nat)) 
   post-condition := post_ledger = ledger 
      && post_admin = admin 
      && post_paused = paused 
      && post_total_supply = total_supply 
      && post_Balance = Balance 
      && transfer_token (0, Amount, c) ;
\end{lstlisting}
\subsubsection{Further entrypoints}
\label{sec:other-entrypoints}

@@

The subsequent property dictates that if \lstinline/getXtzUsdPrice/ is successfully executed, the state must transition to \lstinline/WAITING_FOR_HARBINGER/, and the parameters passed along with the call need to be stored in the \lstinline/clientCallback/ attribute of the storage. This property can be detailed as shown in Listing \ref{lst:kolibri-contract-specification}.

The following property asserts that if \lstinline/getXtzUsdPrice_callback/ is successful, the state should revert to \lstinline/IDLE/, and there must be a transaction transferring the data to the client at the \lstinline/clientCallback/ attribute. This property is specified in Listing \ref{lst:kolibri-contract-specification}.

The final property concerns the entrypoint relation between
\lstinline/getXtzUsdPrice/ and \lstinline/getXtzUsdPrice_callback/ as
depicted in Figure \ref{fig:kolibri-oracle-emtrypoint-relations}. From
the \lstinline/getXtzUsdPrice/ entrypoint, the contract can only be
called with \lstinline/getXtzUsdPrice_callback/ and then
reverted. Invoking the \lstinline/getXtzUsdPrice/ entrypoint after
another call or invoking the \lstinline/getXtzUsdPrice_callback/
entrypoint again is not allowed. This entrypoint relation property is
specified as an entrypoint relation in
Listing~\ref{lst:kolibri-oracle-entrypoint-relations-text}.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.8\textwidth]{kolibri}
    \caption{Kolibri oracle entrypoint relations}
    \label{fig:kolibri-oracle-emtrypoint-relations}
\end{figure}

\begin{lstlisting}[float=tp,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract kolibri = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) | (state = 0)
    post-condition := (state = 1) 
       && (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 
       && Sender = harbinger  
       && s = 'XTZ-USD' 
       && Now - t <= maxDelay 
       && client = Some a 
       && get_contract (nat, a) = Some c    
    post-condition := state = 0 
       && post_client = None 
       && transfer_token(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) 
       && (post_maxDelay = maxDelay) 
       && (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) 
       && (post_client = client) 
       && (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     and (s = 'XTZ-USD') 
     and (Now - t <= maxDelay)  
  | (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  | not (%getXtzUsdRate -> %getXtzUsdRate) 
  | not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}

When verifying the entrypoint \lstinline/setMaxDataDelaySec/, we noticed an unmatched statement. In \cite{kolibri}, the Kolibri oracle is expected to perform two checks on data returned from Harbinger: (1) was the asset returned "XTZ-USD" and (2) was the asset updated in the last 30 minutes. We ran the fail condition check for both the \lstinline/setMaxDataDelaySec/ and \lstinline/getXtzUsdPrice_callback/ entrypoints as shown in Listing \ref{lst:setXtzUsdPrice_callback} and \ref{lst:setMaxDataDelaySec}. Upon examining the failwith conditions, we detected that there is no failwith if the data is more than 30 minutes (equal to 1800 seconds). We can add the predicate \lstinline/maxDelay <= 1800/ in the post condition of the entrypoint \lstinline/setMaxDataDelaySec/, but then the pre and post condition test fails. This means there is no constraint that the data is less than 30 minutes. This constraint is possibly set up by the governor and specified in the \lstinline/MaxDataDelaySec/ field. However, users calling this contract should be aware that it is possible for the governor to set any value (>=0) as desired. This could lead to providing out-of-date rates.
\begin{lstlisting}[float=tp,captionpos=b,caption={Failwith condition for the entrypoint setXtzUsdPrice\_callback},label={lst:setXtzUsdPrice_callback},numbers=left]
1. not (state = 1): 12
2. (state = 1) && not (Sender = harbinger): 3
3. (state = 1) &&  (Sender = harbinger) 
&& not ('XTZUSD' = s): 14
4. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t >= maxDelay): 17
5. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = Some sbvar) 
&& get_contract (Nat, sbvar) = None: Unit
6. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = None): Unit
7. (state = 1) &&  (Sender = harbinger) && Not (Now - t >= 0) 
&& not ('XTZUSD' = s): Unit
\end{lstlisting}

@@

Several tools for automated verification include solc-verify \cite{solc}, VerX \cite{verx}, and Oyente \cite{oyente}. solc-verify processes smart contracts written in Solidity and discharges verification conditions using modular program analysis and SMT solvers. It operates at the level of the contract source code, with properties specified as contract invariants and function pre- and post-conditions provided as annotations in the code by the developer. This approach offers a scalable, automated, and user-friendly formal verification solution for Solidity smart contracts. The core of solc-verify involves translating Solidity contracts to Boogie IVL (Intermediate Verification Language), a simple language designed for verification.

@@

iContract targets the Solidity language. It also utilizes pre and post conditions, similar to our tool, to specify user properties. iContract supports contract invariants, loop invariants, and pre-/post conditions. It locally installs the Solidity compiler to compile a user-provided Solidity file into a JSON file containing the typed abstract syntax tree (AST). Then, iContract analyzes the AST to encode contracts into predicates using the Z3 library. They leverage the NatSpec format to define their own specifications.

>>>>

While designers can create tokens that behave in any way they want, it
is advisable to implement tokens that adhere to established standards.
This adherence guarantees interoperability and compatibility across
different platforms. A token standard defines a set of rules and an
interface that smart contracts must follow to ensure compatibility
with decentralized exchanges, wallets, and other tools within the
ecosystem. For example, ERC-20 \cite{erc} is a token standard used in Ethereum \cite{eth-whitepaper}.

The Tezos blockchain has its own set of standards known as Financial
Applications (FA). Two notable standards are: (1) FA 1.2 \cite{fa}, designed for
fungible tokens, and (2) FA 2 \cite{fas}, a more versatile standard accommodating
both fungible and non-fungible tokens (NFTs). The standard outlines a
smart contract that functions as a ledger, associating identities with
balances. This ledger is designed to facilitate token transfer
operations and enable approvals for spending tokens from other
accounts. Any contract aiming to implement the standard must
incorporate the functions specified in
Listing~\ref{lst:fa12-interface} such as token transfer (transactions
between users), approval (allowing one user to send tokens from
another user's address),  allowance (browsing the amount of tokens
available for an account to sending from another address), user
balance (checking the balance of a user), and current supply (checking
the total supply of tokens across all addresses), where
\lstinline/view a r = pair a (contract r)/ implies an view entrypoint.
A view is an entrypoint that takes an
argument of type \lstinline/a/ and a contract to callback that receives in a value
of type \lstinline/r/. The callback in this case will be another
contract address. Further elaboration on the details of FA 1.2 can be
found in \cite{fatzip}, and we will delve into
these functions in subsequent 
sections.

@@

\subsubsection{transfer}
As illustrated in Listing \ref{lst:fa12-interface}, the
\lstinline/transfer/ entrypoint credits the account associated with
the address provided in the \lstinline/to/ parameter, while debiting
the account corresponding to the \lstinline/from/ parameter with the
specified \lstinline/value/ amount. The FA 1.2 standard mandates that
this call must fail under the following two conditions: (1) In the
event of insufficient balance, this entrypoint MUST fail with the
error \lstinline/NotEnoughBalance/, indicating insufficient funds in
the sender's account to execute the transfer. (2) If the sender's
account has no permission to withdraw the \lstinline/value/ amount of
funds, the call will fail with the \lstinline/NotEnoughAllowance/
error, which also contains the same pair message as the first failure
scenario. To verify these requirements, we utilize the failwith
condition feature of the tool and examine the returned results, as
outlined in Listing~\ref{lst:transfer-udstz-contract}\footnote{There may be some duplicates of the same failwith condition due to the branching of the symbolic interpreter. The results here are shortened. We are working on filtering them in the tool.}.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/transfer/ entrypoint},label={lst:transfer-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (from, ledger) = None : 'NotEnoughBalance'
3. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   sbvar1 < value : 'NotEnoughBalance'
4. not (paused = true) &&
   get_big_map (from, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (Sender, sbvar2) = Some sbvar3 &&
   sbvar3 < value: 'NotEnoughAllowance'
\end{lstlisting}
The fundamental property that this entrypoint should satisfy is the decrement of the allowance of the sender's account by the \lstinline/value/ that has been sent as Listing \ref{lst:specification-transfer}.
\begin{lstlisting}[float,captionpos=b,caption={Specification of the \lstinline/transfer/ entrypoint},label={lst:specification-transfer},numbers=left]
mcontract usdtz = spec 
 storage :=  Pair (ledger: big_map address 
  (pair nat (map address nat))) 
  (Pair (admin: address) 
          (Pair (paused: bool) (total_supply: nat)))
 entrypoint %transfer
   parameter := Pair (from: address)
         (Pair (to: address) (value: nat))
   pre-condition := not (paused  = true)
     && get_big_map (from, ledger) = 
       Some (v: pair nat (map address nat))
     && (v = Pair (n: nat) (m: map address nat))
     && get_map (Sender, m) = Some (k: nat) 
     && (n >= value) && (k >= value)
   post-condition := get_big_map (from, post_ledger) = 
     Some (post_v:  pair nat (map address nat))
     && post_v = Pair (post_n: nat) (post_m: map address nat) 
     && get_map (Sender, post_m) = Some (post_k: nat) 
     && (post_n = n - value) && (post_k = k - value)
\end{lstlisting}

@@

The crucial property to ensure the correctness of the \texttt{approve}
entrypoint is that changing the allowance value from a non-zero value
to another non-zero value is forbidden to prevent a corresponding
attack vector \cite{attack-vector}. The basic idea of this attack is
that when a non-zero value update is allowed, the \lstinline/spender/
can attempt to spend the entire allowance before the new amount is
updated and then spend again with the new value, resulting in a larger
amount than actually allowed. Hence, the contract call must fail when
both the parameter \lstinline/value/ is non-zero and the allowance of
\lstinline/spender/ is also non-zero. We verify this property by
generating a fail condition, so that a non-zero value update is
recorded as a fail condition, as shown in
Listing~\ref{lst:approve-udstz-contract}. This condition indicates
that the implementation ensures the non-zero value update prevention. 
\begin{lstlisting}[float=tp,captionpos=b,caption={Fail conditions for the \lstinline/approve/ entrypoint},label={lst:approve-udstz-contract},numbers=left]
1. (paused = true): 'TokenOperationsArePaused'
2. not (paused = true) &&
   get_big_map (Sender, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (spender, sbvar2) = Some sbvar3 &&
   not (value = 0) &&
   not (sbvar3 = 0): 'UnsafeAllowanceChange'
\end{lstlisting}
If the update is allowed, then the updated value needs to be recorded
correctly in the storage. This property can be divided into two
subproperties, which can be checked separately. Property 1: if the
parameter \lstinline/value/ is zero, then the 
corresponding allowance value of the \lstinline/spender/ from the
owner (which is \SENDER\ in this call) should be zero. This
subproperty is specified as shown in
Listing~\ref{lst:specification-approve-1}. Property 2: if
\lstinline/value/ is non-zero and the allowance of the \lstinline/spender/ is zero, then the \lstinline/value/ should be recorded in the storage after the call. We specify this property using pre- and post-conditions, as illustrated in Listing~\ref{lst:specification-approve-2}.

\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/approve/ entrypoint (Property 1)},label={lst:specification-approve-1},numbers=left]
entrpoint %approve
   parameter := Pair (spender: address) (value: nat)
   pre-condition := not (paused = true)
      && get_big_map (Sender, ledger) = 
        Some (v: pair nat (map address nat))
      && v = Pair (n: nat) (m: map address nat) 
      && get_map (spender, m) = Some (k: nat) 
      && not (k = 0) | (value = 0)
   post-condition := get_big_map (Sender, post_ledger) = 
        Some (post_v:  (pair nat (map address nat))
      && post_v = Pair (post_n: nat) (post_m: map address nat) 
      && get_map (spender, post_m) = Some (post_k: nat) 
      && (post_k = 0) ;
\end{lstlisting}
\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/approve/ entrypoint (Property 2)},label={lst:specification-approve-2},numbers=left]
   pre-condition := not (paused = true) 
      && get_big_map (Sender, ledger) = 
        Some (v: option (pair nat (map address nat))) 
      && v = Pair (n: nat) (m: map address nat) 
      && get_map (spender, m) = Some (k: nat) 
      && (k = 0) | not (value = 0)
   post-condition := get_big_map (Sender, post_ledger) = 
        Some (post_v: option (pair nat (map address nat))) 
      && post_v = Pair (post_n: nat) (post_m: map address nat) 
      && get_map (spender, post_m) = Some (post_k: nat) 
      && (post_k = value) ;
\end{lstlisting}

@@

As an example, we demonstrate how to verify the \lstinline/getBalance/
entrypoint,  shown in Listing~\ref{lst:specification-get-balance}. The
other entrypoints \lstinline/getAllowance/ and
\lstinline/getTotalSupply/ can be specified similarly.  There are two
cases to consider: (1) if the \lstinline/owner/ is in the
\lstinline/ledger/ big map and (2) if it is not. For the former case,
lines 5-8 specify that \lstinline/owner/ is in the \lstinline/ledger/
big map, as the \lstinline/get_big_map/ function returns a value
\lstinline/Some value/ of the \TOPTION\ type, and this
\lstinline/value/ is a pair of a natural number \lstinline/k/ and a
map \lstinline/m/. The postcondition guarantees that the storage
remains unchanged, as specified in lines 10-13, and the call returns a
\lstinline/TRANSFER_TOKENS/ operation with the \lstinline/k/ number
and \lstinline/Amount/ to be passed to address \lstinline/c/. The
precondition in lines 16-17 specifies the latter case when the
\lstinline/owner/ does not exist in the big map \lstinline/ledger/; in
this case, the number \lstinline/0/ should be sent, and again, the
storage should remain unchanged.

\begin{lstlisting}[float=tp,captionpos=b,caption={Specification of the \lstinline/getBalance/ entrypoint},label={lst:specification-get-balance},numbers=left]
entrypoint %getBalance
   parameter := Pair (owner: address) (c: contract nat)
   pre-condition :=  get_big_map (owner, ledger) = 
        Some v: option (pair nat (map address nat)) 
      && v  = Pair (k: nat) (m: (map address nat))
   post-condition := post_ledger = ledger 
      && post_admin = admin 
      && post_paused = paused 
      && post_total_supply = total_supply 
      && post_Balance = pre_Balance 
      && transfer_token (k, Amount, c) ;

   pre-condition :=  get_big_map (owner, ledger) = 
        None option (pair nat (map address nat)) 
   post-condition := post_ledger = ledger 
      && post_admin = admin 
      && post_paused = paused 
      && post_total_supply = total_supply 
      && post_Balance = pre_Balance 
      && transfer_token (0, Amount, c) ;
\end{lstlisting}
\subsubsection{Further entrypoints}
\label{sec:other-entrypoints}

@@

The subsequent property dictates that if \lstinline/getXtzUsdPrice/ is successfully executed, the state must transition to \lstinline/WAITING_FOR_HARBINGER/, and the parameters passed along with the call need to be stored in the \lstinline/clientCallback/ attribute of the storage.

The following property asserts that if \lstinline/getXtzUsdPrice_callback/ is successful, the state should revert to \lstinline/IDLE/, and there must be a transaction transferring the data to the client at the \lstinline/clientCallback/ attribute.

The final property concerns the entrypoint relation between
\lstinline/getXtzUsdPrice/ and \lstinline/getXtzUsdPrice_callback/ as
depicted in Figure \ref{fig:kolibri-oracle-emtrypoint-relations}. From
the \lstinline/getXtzUsdPrice/ entrypoint, the contract can only be
called with \lstinline/getXtzUsdPrice_callback/ and then
reverted. Invoking the \lstinline/getXtzUsdPrice/ entrypoint after
another call or invoking the \lstinline/getXtzUsdPrice_callback/
entrypoint again is not allowed.

The specification of  the contract is showned in Listing \ref{lst:kolibri-contract-specification}. Here, we shortened the names of the variables for the storage.
\begin{figure}[tp]
    \centering
    \includegraphics[width=0.8\textwidth]{kolibri}
    \caption{Kolibri oracle entrypoint relations}
    \label{fig:kolibri-oracle-emtrypoint-relations}
\end{figure}

\begin{lstlisting}[float=tp,captionpos=b,caption={Kolibri oracle contract specification},label={lst:kolibri-contract-specification},numbers=left]
mcontract kolibri = spec 
  storage := (Pair (Pair (client: option address) 
     (governor: address))
     (Pair (harbinger: address) 
           (Pair (maxDelay: nat) (state: int))))
  entrypoint %getXtzUsdRate
    code := {...}
    parameter := (c : contract nat) 
    pre-condition := (Amount = 0) | (state = 0)
    post-condition := (state = 1) 
       && (client = Some Sender) ;
  entrypoint %getXtzUsdRate_callback
    code := {...}
    parameter := Pair (s: string) 
       (Pair (t: timestamp) (n: nat))
    pre-condition :=  state = 1 
       && Sender = harbinger  
       && s = 'XTZ-USD' 
       && Now - t <= maxDelay 
       && client = Some a 
       && get_contract (nat, a) = Some c    
    post-condition := state = 0 
       && post_client = None 
       && transfer_token(n, 0, c) ;
  entrypoint %setGovernorContract
    code := {...}
    parameter := (gover : address) 
    pre-condition :=  (Sender = governor)                  
    post-condition := (governor = gover) 
       && (post_maxDelay = maxDelay) 
       && (post_harbinger = harbinger) ;
  entrypoint %setMaxDataDelaySec
    code := {...}
    parameter := (delay: nat)
    pre-condition := (Sender = governor)             
    post-condition := (maxDelay = delay) 
       && (post_client = client) 
       && (post_harbinger = harbinger) ;
  (%getXtzUsdRate -> %getXtzUsdRate_callback) with
     (Sender = harbinger)
     and (s = 'XTZ-USD') 
     and (Now - t <= maxDelay)  
  | (%getXtzUsdRate_callback  -> %getXtzUsdRate) with 
    (Amount = 0) 
  | not (%getXtzUsdRate -> %getXtzUsdRate) 
  | not (%getXtzUsdRate_callback -> %getXtzUsdRate_callback)
\end{lstlisting}
When verifying the entrypoint \lstinline/setMaxDataDelaySec/, we noticed an unmatched statement. In \cite{kolibri}, the Kolibri oracle is expected to perform two checks on data returned from Harbinger: (1) was the asset returned "XTZ-USD" and (2) was the asset updated in the last 30 minutes. We ran the fail condition check for both the \lstinline/setMaxDataDelaySec/ and \lstinline/getXtzUsdPrice_callback/ entrypoints as shown in Listing \ref{lst:setXtzUsdPrice_callback} and \ref{lst:setMaxDataDelaySec}. Upon examining the failwith conditions, we detected that there is no failwith if the data is more than 30 minutes (equal to 1800 seconds). We can add the predicate \lstinline/maxDelay <= 1800/ in the post condition of the entrypoint \lstinline/setMaxDataDelaySec/, but then the pre and post condition test fails. This means there is no constraint that the data is less than 30 minutes. This constraint is possibly set up by the governor and specified in the \lstinline/MaxDataDelaySec/ field. However, users calling this contract should be aware that it is possible for the governor to set any value (>=0) as desired. This could lead to providing out-of-date rates.
\begin{lstlisting}[float=tp,captionpos=b,caption={Failwith condition for the entrypoint setXtzUsdPrice\_callback},label={lst:setXtzUsdPrice_callback},numbers=left]
1. not (state = 1): 12
2. (state = 1) && not (Sender = harbinger): 3
3. (state = 1) &&  (Sender = harbinger) 
&& not ('XTZUSD' = s): 14
4. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t >= maxDelay): 17
5. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = Some sbvar) 
&& get_contract (Nat, sbvar) = None: Unit
6. (state = 1) &&  (Sender = harbinger) && ('XTZUSD' = s)
&& (Now - t >= 0) && (Now - t < maxDelay) 
&& (client = None): Unit
7. (state = 1) &&  (Sender = harbinger) && Not (Now - t >= 0) 
&& not ('XTZUSD' = s): Unit
\end{lstlisting}

@@

There are exsiting tools for automated verification include solc-verify \cite{solc}, VerX \cite{verx}, and Oyente \cite{oyente}. solc-verify processes smart contracts written in Solidity and discharges verification conditions using modular program analysis and SMT solvers. It operates at the level of the contract source code, with properties specified as contract invariants and function pre- and post-conditions provided as annotations in the code by the developer. This approach offers a scalable, automated, and user-friendly formal verification solution for Solidity smart contracts. The core of solc-verify involves translating Solidity contracts to Boogie IVL (Intermediate Verification Language), a simple language designed for verification.

@@

iContract targets the Solidity language. It also utilizes pre and post conditions, similar to our tool, to specify user properties. It locally installs the Solidity compiler to compile a user-provided Solidity file into a JSON file containing the typed abstract syntax tree (AST). Then, iContract analyzes the AST to encode contracts into predicates using the Z3 library. They leverage the NatSpec format to define their own specifications.
